---
title: "LV_analysis"
author: "Carlos Olivares"
date: "3/7/2023"
output: html_document
---

```{r echo=FALSE}
library(tidyverse)
library(deSolve)
library("RColorBrewer")
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## One Family equation per taxa
Considering only one family that not interact with other bacterial families following the equation

$$ \frac{d}{dt} N_{k} = \alpha_k N_{k}  - \mu_{k,k} N_{k}^2 $$
This equation shows a competition between a linear and a quadratic term and considering $\alpha_k = 20$ and $\mu_{k,k} = 2 $ and $N_{k}^{*} = 10$ in growth rate, 
```{r}
ggplot() + 
  geom_function(fun= function(x) 20*x) + 
  geom_function(fun= function(x) 2*x^2) + 
  theme_bw() +
  xlim(0, 20)
```

in which the fix point is $N_{k}^{*} = \alpha_{k}/ \mu_{k,k}$. 

Particularly this equation has an analytical solution

$$ N_{k} = \frac{ \alpha_{k} N_{k}^{0} \exp(\alpha t) }{\alpha_k + \mu_{k,k} N_{k}^{0}  \left[\exp(\alpha_{k} t) - 1 \right] } $$
Considering the same fix point $N_{k}^{*} $ but with 2 different initial conditions we get the following behavoir

```{r}
f_Nk_sol <- function(t, Nk0, alpha_k, mu_kk){
  return( alpha_k*Nk0*exp(alpha_k*t) / (alpha_k + mu_kk*Nk0*(exp(alpha_k*t) -1 ) ) )
}

colores_list <- brewer.pal(n = 8, name = "RdBu")
ggplot() +
  geom_function(fun=f_Nk_sol, args=list(Nk0=150, alpha_k=400, mu_kk=4), col=colores_list[1]) +
  geom_function(fun=f_Nk_sol, args=list(Nk0=150, alpha_k=200, mu_kk=2), col=colores_list[2]) +
  geom_function(fun=f_Nk_sol, args=list(Nk0=150, alpha_k=100, mu_kk=1), col=colores_list[3]) +
  geom_function(fun=f_Nk_sol, args=list(Nk0= 50, alpha_k=400, mu_kk=4), col=colores_list[4]) +
  geom_function(fun=f_Nk_sol, args=list(Nk0= 50, alpha_k=200, mu_kk=2), col=colores_list[5]) +
  geom_function(fun=f_Nk_sol, args=list(Nk0= 50, alpha_k=100, mu_kk=1), col=colores_list[6]) +
  geom_function(fun=f_Nk_sol, args=list(Nk0=150, alpha_k=1,  mu_kk=0.01), col=colores_list[7]) +
  theme_bw()
```
hola hola
```{r}
f_Nk_sol_02 <- function(t, Nk0, alpha_k, mu_kk){
  return( alpha_k*Nk0 / (alpha_k*exp(-alpha_k*t) + mu_kk*Nk0*(1 - exp(-alpha_k*t) ) ) )
}


colores_list <- brewer.pal(n = 8, name = "RdBu")
ggplot() +
  geom_function(fun=f_Nk_sol_02, args=list(Nk0=150, alpha_k=1000, mu_kk=10), col=colores_list[1]) +
  geom_function(fun=f_Nk_sol_02, args=list(Nk0=150, alpha_k=200, mu_kk=2), col=colores_list[2]) +
  geom_function(fun=f_Nk_sol_02, args=list(Nk0=150, alpha_k=100, mu_kk=1), col=colores_list[3]) +
  geom_function(fun=f_Nk_sol_02, args=list(Nk0= 50, alpha_k=400, mu_kk=4), col=colores_list[4]) +
  geom_function(fun=f_Nk_sol_02, args=list(Nk0= 50, alpha_k=200, mu_kk=2), col=colores_list[5]) +
  geom_function(fun=f_Nk_sol_02, args=list(Nk0= 50, alpha_k=100, mu_kk=1), col=colores_list[6]) +
  geom_function(fun=f_Nk_sol_02, args=list(Nk0=150, alpha_k=1,  mu_kk=0.01), col=colores_list[7]) +
  theme_bw()
```

Parameters to show
```{r, echo=FALSE}
Nk_fp <- 1e5
tbl_tmp <- tibble(
  alpha_k = c(100, 10.0, 1.00, 1e-1,   1e-2)
  # mu_kk   = c(  1,  0.1, 0.01, 1e-3, 1e-4)
  ) %>% mutate(
    mu_kk = alpha_k / Nk_fp
  )

tbl_params <- rbind( tbl_tmp %>% mutate(Nk0=1.5e5), tbl_tmp %>% mutate(Nk0=0.5e5) )
tbl_params
```


```{r, echo=FALSE}
f_Nk_sol <- function(t, Nk0, alpha_k, mu_kk){
  return( alpha_k*Nk0*exp(alpha_k*t) / (alpha_k + mu_kk*Nk0*(exp(alpha_k*t) -1 ) ) )
}

f_Nk_sol_02 <- function(t, Nk0, alpha_k, mu_kk){
  
  return( alpha_k*Nk0 / (alpha_k*exp(-alpha_k*t) + mu_kk*Nk0*(1 - exp(-alpha_k*t) ) ) )
}

tt <- seq(0, 200, 0.1)

tbl_analytical <- tbl_params %>% 
  group_by(alpha_k, mu_kk, Nk0) %>% 
  mutate(
    data = list(
      data.frame(
      time = tt,
      Nk = f_Nk_sol_02(tt, Nk0, alpha_k, mu_kk)
      ) 
    ) ) %>% 
  unnest(cols = c("data"))
tbl_analytical

```

```{r, echo=FALSE}
p <- ggplot(data=tbl_analytical, aes(x=time, y=Nk, color=as.factor(alpha_k) )) + 
  geom_line() + 
  theme_bw() +
  facet_wrap(~Nk0)
p
```

Considering only one family we get the 

$$ \frac{d}{dt} N_{k} = \alpha_k N_{k}  - \mu_{k,k} N_{k}^2 $$
Now show how this change under different parameters


```{r}
tbl_params
```


```{r}

f_dNkdt_1Fam_NoAb <- function(t, Nk, parameters){
  with(
    as.list(c(Nk, parameters)), {
      dNkdt <- alpha_k*Nk - mu_k_k*Nk*Nk
      list(dNkdt)
    }
  )
}

Nk0 <- 1.5*Nk_fp
times <- seq(0, 100, by = 0.1)

alpha_k <- 1e-1
mu_k_k  <- alpha_k/Nk_fp

ini_cond <- c(Nk=Nk0)

parameters <- list(alpha_k=alpha_k, mu_k_k=mu_k_k)

out <- ode(y = ini_cond, times = times,
           func = f_dNkdt_1Fam_NoAb, parms = parameters,
           method = "ode45")
# out
data_Nk <- tibble(time = out[,1], Nk_t = out[,2] )
ggplot(data=data_Nk, aes(x=time, y=Nk_t)) + geom_line()
```




```{r}

f_dNkdt_1Fam_NoAb <- function(t, Nk, parameters){
  with(
    as.list(c(Nk, parameters)), {
      dNkdt <- alpha_k*Nk - mu_k_k*Nk*Nk
      list(dNkdt)
    }
  )
}

f_Nk_num_1Fam_NoAb <- function(times, Nk0, alpha_k, mu_k_k){
  ini_cond <- c(Nk=Nk0)
  parameters <- list(alpha_k=alpha_k, mu_k_k=mu_k_k)
  out <- ode(y = ini_cond, times = times,
             func = f_dNkdt_1Fam_NoAb, parms = parameters,
             method = "ode45")
  data_Nk <- tibble(time = out[,1], Nk = out[,2] )
  return(data_Nk)
}

# Nk0 <- 2e5
# times <- seq(0, 100, by = 0.1)
# 
# alpha_k <- 1e-1
# mu_k_k  <- 1e-6

f_Nk_num_1Fam_NoAb(tt, 2e5, 1e-1, 1e-6)

tbl_numeric <- tbl_params %>% 
  group_by(alpha_k, mu_kk, Nk0) %>% 
  mutate(
    data = list(
      f_Nk_num_1Fam_NoAb(tt, Nk0, alpha_k, mu_kk)
    ) ) %>% 
  unnest(cols = c("data"))
tbl_numeric


```

```{r}
p <- ggplot(data=tbl_numeric, aes(x=time, y=Nk, color=as.factor(alpha_k) )) + 
  geom_line() + 
  theme_bw() +
  facet_wrap(~Nk0)
p

```

## One Family with Interaction with Antibiotic

```{r}

f_dNkdt_1Fam_Ab_lin_Cf0 <- function(t, Nk, parameters){
  with(
    as.list(c(Nk, parameters)), {
      dNkdt <- alpha_k*Nk - mu_k_k*Nk*Nk - gamma_k*Cf0*Nk
      list(dNkdt)
    }
  )
}


f_Nk_num_1Fam_NoAb <- function(times, Nk0, alpha_k, mu_k_k, gamma_k, Cf0){
  ini_cond <- c(Nk=Nk0)
  parameters <- list(alpha_k=alpha_k, mu_k_k=mu_k_k)
  out <- ode(y = ini_cond, times = times,
             func = f_dNkdt_1Fam_Ab_lin_Cf0, parms = parameters,
             method = "ode45")
  data_Nk <- tibble(time = out[,1], Nk = out[,2] )
  return(data_Nk)
}

```


```{r}
ve <- function(x) {
  exp(-0.03*log(10)*x)
}
ggplot() + 
  geom_function(fun=ve) + 
  xlim(0, 150) + 
  scale_y_log10()
```

