---
title: "Lotka Volterra Simulations for 7 particle system Jinju's inference"
author: "Carlos Olivares"
date: "10/3/2024"
output:
  html_document:
    code_folding: hide
    toc: TRUE
    number_sections: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, cache=TRUE, collapse = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
theme_ceor <-theme_bw() +
  # theme(legend.position = "none")+
  theme(
    plot.title = element_text(size = 24,hjust=-0.03),
    axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text=element_text(size=16),
    axis.title=element_text(size=18),
    axis.line.x = element_line(size = 0.5, linetype = "solid", colour = "black"),
    axis.line.y = element_line(size = 0.5, linetype = "solid", colour = "black")
  )


```



# Objective

Make simulations of a lotka-volterra system, considering for N species.
 - Analyze the effect on the number of interactions
 - Stability issues
 - Take already inferred parameters from other papers for a stability check.
 - I want to explore the stability and also what is the temporal evolution of the species.
 - 

# Lotka-Volterra


```{r}
# anscombe
# anscombe %>%
#   pivot_longer(
#     everything(),
#     cols_vary = "slowest",
#     names_to = c(".value", "set"),
#     names_pattern = "(.)(.)"
#   )

```




## Interaction Matrix

```{r}
# install.packages("reshape2")
# Load the ggplot2 package
library(ggplot2)
library(reshape2)

# Example matrix
# mat <- matrix(runif(100), nrow = 10, ncol = 10)

fln_interactions_mat <- "c_001_deterministic/interaction_matrix_N_7.mat"

# mat_interactions <- read.csv(fln_interactions_mat, header = FALSE) %>% 
#   as.matrix()

mat_interactions <- read.csv(fln_interactions_mat, header = FALSE) %>% as_tibble(.name_repair = ~as.character(seq_along(.))) %>% as.matrix()
colnames(mat_interactions) <- NULL


# Convert matrix to data frame
mat_df <- melt(mat_interactions)

# Plot using ggplot2
lbl_title <- "" # "Interaction Matrix"
lbl_x <- "Species"
lbl_y <- "Species"
ggplot(mat_df, aes(Var2, Var1, fill = value)) +
  geom_tile(color = "white") +  # Plot tiles
  geom_text(aes(label = round(value, 1)), color = "black") +  # Add value text
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = lbl_title, x = lbl_x, y = lbl_y) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_fixed() +
  scale_y_reverse() +  # Flip the y-axis for intuitive plotting
  theme_ceor
```



## Species Parameters

```{r}
fln_species <- "GLV_Simulations/SimpleGLV/species_N_3.csv"
df_species <- read.csv(fln_species)
v_tmp <- df_species %>% pull(AbsAbun) 
-mat_interactions %*% v_tmp
```

```{r}
t(mat_interactions) %*% v_tmp

```


```{r}
mat_interactions
```


```{r}

eigen_mat_interactions <- eigen(mat_interactions)

# eigen_mat_interactions$vectors[,1]**2 %>% sum()
```


## Eigen values


```{r}

# Calculate eigenvectors
eigen_mat_interactions <- eigen(mat_interactions)
eigen_mat_interactions_vectors <- eigen_mat_interactions$vectors

# Prepare data for plotting
# Create a data frame with origin (0, 0) and the end points of the eigenvectors
unitary_vectors <- data.frame(
  x_start = rep(0, ncol(eigen_mat_interactions_vectors)),  # Start x-coordinates (origin)
  y_start = rep(0, ncol(eigen_mat_interactions_vectors)),  # Start y-coordinates (origin)
  x_end = eigen_mat_interactions_vectors[1, ],              # End x-coordinates (eigenvectors)
  y_end = eigen_mat_interactions_vectors[2, ]               # End y-coordinates (eigenvectors)
)

r <- 1  # radius
theta <- seq(0, 2 * pi, length.out = 100)  # angle from 0 to 2*pi
circle_data <- data.frame(
    x = r * cos(theta),
    y = r * sin(theta)
  )

lbl_title <- ""
lbl_x <- ""
lbl_y <- ""
# Plot using ggplot2
ggplot(unitary_vectors) +
  geom_segment(aes(x = x_start, y = y_start, xend = x_end, yend = y_end), 
               arrow = arrow(length = unit(0.2, "inches")), color = "black") +
  geom_path(data=circle_data, aes(x = x, y = y), color = "blue", alpha=0.5, linetype='dashed', size = 1) +
  # coord_cartesian(
  #   
  # ) +
  labs(title = lbl_title, x = lbl_x, y = lbl_x) +
  geom_text(aes(x = x_end, y = y_end, label = paste0("v", 1:ncol(eigen_mat_interactions_vectors))), vjust = -1.5) +
  coord_fixed() + 
  theme_ceor
```



## Trajectories

```{r, fig.height=6, fig.width=8}
df_trajectories <- read.csv("GLV_Simulations/SimpleGLV/df_trajectories_N_3.csv")
df_trajectories %>% 
  pivot_longer(cols = contains("Species"), names_to = "id_species", values_to = "AbsAbun") %>% 
  dplyr::mutate(id_species = str_replace_all(id_species, "Species.", "")) %>% 
  dplyr::mutate(id_species = as.integer(id_species)) %>% 
  ggplot() +
  geom_line(aes(x=Time, y=AbsAbun, color=id_species, group=id_species)) +
  theme_ceor

```



## Diversity


```{r}
df_diversity <- read.csv("GLV_Simulations/SimpleGLV/df_diversity_N_3.csv")
df_diversity %>% 
  ggplot(aes(x=time, y=Shannon))+
  # geom_line(data=DF_DIVERSITY, aes(x=V1, y=V3)) +
  geom_line() +
  # geom_line(aes(y=V2), color = 'red') + 
  # geom_point() +
  coord_cartesian(
    x=c(0,100)
  ) +
  # scale_y_log10()  +
  theme_ceor
```



```{r, fig.height=8, fig.width=16}
# # fln_csv <- "GLV_Simulations/output.csv"
# fln_csv <- "GLV_Simulations/SimpleGLV/data.csv"
# 
# df_simul <- read.csv(fln_csv, sep = '\t')
# df_species_AbsAbun <- df_simul %>% 
#   dplyr::select(-X) %>% 
#   pivot_longer(
#     cols = contains("Species."),
#     names_to = "Species",
#     values_to = "Abundance"
#     # cols_vary = "slowest",
#     # names_to = c(".value", "set"),
#     # names_pattern = "(.)(.)"
#   ) %>% 
#   dplyr::mutate(Species = as.integer( str_replace_all(Species, "Species.", "") ) )
#   # head(100) %>%
# 
# 
# df_species_AbsAbun %>%
#   group_by(Time) %>% 
#   summarise(
#     
#   )
# 
# plt_tmp <- df_species_AbsAbun %>% 
#   ggplot(aes(x=Time, y=Abundance, color=as.factor(Species) ) ) +
#   geom_line(data=DF_DIVERSITY, aes(x=V1, y=V3), inherit.aes = FALSE, color='black') +
#   geom_line() +
#   # geom_point() +
#   scale_y_log10() +
#   coord_cartesian(
#     xlim = c(0, 100)
#   ) +
#   theme_ceor
# 
# plotly::ggplotly(plt_tmp)

```









